---
title: "Untitled"
author: "Tom Swan"
date: "08/07/2021"
output: html_document
---

R script associated with distance based redundancy (dbRDAs) analyses for Aedes albopictus in the Torres Strait Islands. Models tested for an association between aircraft connectivity and genetic differentiation (pairwise Fst), conditioned by geographic distance. 


```{r}
library(tidyverse)
library(psych)
library(vegan)
library("vcfR")
library("vegan")
library("poppr")
library("pegas")
library("ape")
library("adegenet")
library("ade4")
library("MASS")
library("PBSmapping")
library("plotly")
library("rgdal")
library("installr")
library("sp")
library("genepop")
library("geosphere")
library("tidyverse")
library("dplyr")
library("plotly")
library(ggplot2)
library(reshape)
library(ggpubr)
```

1. Import the Fst scores

```{r}
Fst_TSI_full <- read_csv(file = 'FST_scores_full.csv')

#checking the dimensions. Yes this is the matrix with all islands
dim(Fst_TSI_full)

```

2. convert the Fst scores to 'dist' format

```{r}
dist_Fst_TSI_full <- dist(Fst_TSI_full)
str(dist_Fst_TSI_full)
```

3. Import the geographic coordinates for each island (just one pair of coorindates per island)

```{r}

TSI_22_whole_pops <- read.delim("tsi22_whole_pops.txt", header = F)

str(TSI_22_whole_pops)

row.names(TSI_22_whole_pops) <- TSI_22_whole_pops$V1
TSI_22_whole_pops_rows <- TSI_22_whole_pops[,-1]

#column binding the coordinates together

TSI_22_whole_pops_rows <- cbind(TSI_22_whole_pops_rows$V2, TSI_22_whole_pops_rows$V3)

str(TSI_22_whole_pops_rows)
```
4. Calculating the distance between coordinates (distm function) and converting this to 'dist' format

```{r}
#calculating the distance between points - regular distance
dist_TSI_22_whole_pops_rows <- distm(TSI_22_whole_pops_rows)
str(dist_TSI_22_whole_pops_rows)
plot(dist_TSI_22_whole_pops_rows)

#now converting this to 'dist' format
dist_dist_TSI_22_whole_pops_rows <- dist(dist_TSI_22_whole_pops_rows)
  str(dist_dist_TSI_22_whole_pops_rows)

#converting this to dist

#dist_TSI_22_whole_pops <- dist(TSI_22_whole_pops)
#str(dist_TSI_22_whole_pops)
#plot(dist_TSI_22_whole_pops)

#str(dist_TSI_22_whole_pops)

#USIND the distm values, converting this to dist

```
4. Converting the geographic distance to principle components to use in the dbRDA

```{r}
#TSI_geo_distance_pcs <- pcnm(dist_TSI_22_whole_pops_rows)

#str(TSI_geo_distance_pcs)

#using the distm to convert to PCs

distm_TSI_geo_dist_pcs <- pcnm(dist_dist_TSI_22_whole_pops_rows)
str(distm_TSI_geo_dist_pcs)


```
5. running the dbRDA with the distm for the geographical system

```{r}
#genetic is: dist_Fst_TSI_full
#geographic is: distm_TSI_geo_dist_pcs

distm_draft_TSI_genetic_geographic_with_pcs <- capscale(dist_Fst_TSI_full ~ distm_TSI_geo_dist_pcs$vectors[,1] + distm_TSI_geo_dist_pcs$vectors[,2] + distm_TSI_geo_dist_pcs$vectors[,3] + distm_TSI_geo_dist_pcs$vectors[,4] + distm_TSI_geo_dist_pcs$vectors[,5] + distm_TSI_geo_dist_pcs$vectors[,6] + distm_TSI_geo_dist_pcs$vectors[,7])

set.seed(5)

distm_genetic_geo <- anova.cca(distm_draft_TSI_genetic_geographic_with_pcs, by = "margin", permutations = 9999)
#the same result as above

p.adjust(distm_genetic_geo$`Pr(>F)`, method = "BH")

#PCs 1 and 2 are suitable to condition the model. 

#Not used - overall model testing all vectors with Fst:

fst_geo_dist <- capscale(dist_Fst_TSI_full ~ distm_TSI_geo_dist_pcs$vectors)

set.seed(5)

fst_geo_dist_anova <- anova.cca(fst_geo_dist, by = "margin", permutations = 9999)

```

Running a dbRDA for the number of # plane landings on each island over a given year (July 2020 - July 2021)

6. Import the plane data

```{r}
plane_landings_full <- read.csv(file = "plane_landings_full.csv")

```

7. Convert plane landings into 'dist'

```{r}
plane_landings_full_dist <- dist(plane_landings_full)
str(plane_landings_full_dist)
```

8. Convert plane landings into PCs

```{r}
plane_landings_full_pcs <- pcnm(plane_landings_full_dist)
```

9. Run the dbRDA with genetic, plane landings full, conditioned by geo distance

```{r}
#this is the best model.
dbRDAplane_landings <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors + Condition(distm_TSI_geo_dist_pcs$vectors[,c(1,2)]))

set.seed(5)

anova.cca(dbRDAplane_landings, by = "margin", permutations = 9999)

#This is no where near close to significance. How about if I remove the geographic distance conditioning?

#try to get the sum of squares for just geo dist
                       
#with removing the conditioning by the geo
no_geo_dbRDAplane_landings <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors)

set.seed(5)

anova.cca(no_geo_dbRDAplane_landings, by = "margin", permutations = 9999)

#removing the conditioning by the geo results in this being highly insignificant. 

```
10. Condition the model by population size

10a: Import the island population size
```{r}
#import the island population size

island_population <- read.csv("island_population.csv")
```

11: Convert this island population to 'dist' format

```{r}
island_population_dist <- dist(island_population)
str(island_population_dist)
```

12: Convert this 'dist' format to PCs

```{r}
island_population_pcs <- pcnm(island_population_dist)

```

13. Run the Fst against all the PCs for the population size

```{r}
distm_draft_TSI_genetic_population_with_pcs <- capscale(dist_Fst_TSI_full ~ island_population_pcs$vectors[,1] + island_population_pcs$vectors[,2] + island_population_pcs$vectors[,3] + island_population_pcs$vectors[,4] + island_population_pcs$vectors[,5] + island_population_pcs$vectors[,6] + island_population_pcs$vectors[,7] + island_population_pcs$vectors[,8]+ island_population_pcs$vectors[,9] + island_population_pcs$vectors[,10])

set.seed(5)

distm_genetic_populations <- anova.cca(distm_draft_TSI_genetic_population_with_pcs, by = "margin", permutations = 9999)

distm_genetic_populations

#it looks like they're 3 PCs which are significant. 

p.adjust(distm_genetic_populations$`Pr(>F)`, method = "BH")

#with the correction, there is only two PCs which are significant. That being, PC 3 and PC 9

```

14. Run the dbRDA with genetic ~ plane landings full, conditioned by both geo distance and population size

```{r}
#this is the best model.
dbRDAplane_landings_geo_pop_conditions <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors + Condition(distm_TSI_geo_dist_pcs$vectors[,c(1,2)],island_population_pcs$vectors[,c(3,9)]), distance = "Euclidean")

set.seed(5)

anova.cca(dbRDAplane_landings_geo_pop_conditions, by = "margin", permutations = 9999)

dbRDAplane_landings_geo_pop_conditions <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors + Condition(distm_TSI_geo_dist_pcs$vectors[,c(1,2)]))

set.seed(5)

anova.cca(dbRDAplane_landings_geo_pop_conditions, by = "margin", permutations = 9999)


#interesting - this is close to significance. How about if I remove the geographic distance conditioning?

#try to get the sum of squares for just geo dist
                       
#with removing the conditioning by the geo
no_geo_dbRDAplane_landings <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors)

set.seed(5)

anova.cca(no_geo_dbRDAplane_landings, by = "margin", permutations = 9999)

#removing the conditioning by the geo results in this being highly insignificant. 

#Conditioning for both significant PCs Geography and Population Size

set.seed(5)

anova.cca(two_conditions_dbRDAplane_landings, by = "margin", permutations = 9999)


```


