---
title: "Untitled"
author: "Tom Swan"
date: "08/07/2021"
output: html_document
---

R script associated with distance based redundancy (dbRDAs) analyses for Aedes albopictus in the Torres Strait Islands. Models tested for an association between aircraft connectivity and genetic differentiation (pairwise Fst), conditioned by geographic distance. 


```{r}
library(tidyverse)
library(psych)
library(vegan)
library("vcfR")
library("vegan")
library("poppr")
library("pegas")
library("ape")
library("adegenet")
library("ade4")
library("MASS")
library("PBSmapping")
library("plotly")
library("rgdal")
library("installr")
library("sp")
library("genepop")
library("geosphere")
library("tidyverse")
library("dplyr")
library("plotly")
library(ggplot2)
library(reshape)
library(ggpubr)
library("vcfR")

```

1. Import the Fst scores

```{r}
Fst_TSI_full <- read_csv(file = 'FST_scores_full.csv')

#checking the dimensions. Yes this is the matrix with all islands
dim(Fst_TSI_full)

```

2. convert the Fst scores to 'dist' format

```{r}
dist_Fst_TSI_full <- dist(Fst_TSI_full)
str(dist_Fst_TSI_full)
```

3. Import the geographic coordinates for each island (just one pair of coorindates per island)

```{r}

TSI_22_whole_pops <- read.delim("tsi22_whole_pops.txt", header = F)

str(TSI_22_whole_pops)

row.names(TSI_22_whole_pops) <- TSI_22_whole_pops$V1
TSI_22_whole_pops_rows <- TSI_22_whole_pops[,-1]

#column binding the coordinates together

TSI_22_whole_pops_rows <- cbind(TSI_22_whole_pops_rows$V2, TSI_22_whole_pops_rows$V3)

str(TSI_22_whole_pops_rows)
```
4. Calculating the distance between coordinates (distm function) and converting this to 'dist' format

```{r}
#calculating the distance between points - regular distance
dist_TSI_22_whole_pops_rows <- distm(TSI_22_whole_pops_rows)
str(dist_TSI_22_whole_pops_rows)
plot(dist_TSI_22_whole_pops_rows)

#now converting this to 'dist' format
dist_dist_TSI_22_whole_pops_rows <- dist(dist_TSI_22_whole_pops_rows)
  str(dist_dist_TSI_22_whole_pops_rows)

#converting this to dist

#dist_TSI_22_whole_pops <- dist(TSI_22_whole_pops)
#str(dist_TSI_22_whole_pops)
#plot(dist_TSI_22_whole_pops)

#str(dist_TSI_22_whole_pops)

#USIND the distm values, converting this to dist

```
4. Converting the geographic distance to principle components to use in the dbRDA

```{r}
#TSI_geo_distance_pcs <- pcnm(dist_TSI_22_whole_pops_rows)

#str(TSI_geo_distance_pcs)

#using the distm to convert to PCs

distm_TSI_geo_dist_pcs <- pcnm(dist_dist_TSI_22_whole_pops_rows)
str(distm_TSI_geo_dist_pcs)


```
5. running the dbRDA with the distm for the geographical system

```{r}
#genetic is: dist_Fst_TSI_full
#geographic is: distm_TSI_geo_dist_pcs

distm_draft_TSI_genetic_geographic_with_pcs <- capscale(dist_Fst_TSI_full ~ distm_TSI_geo_dist_pcs$vectors[,1] + distm_TSI_geo_dist_pcs$vectors[,2] + distm_TSI_geo_dist_pcs$vectors[,3] + distm_TSI_geo_dist_pcs$vectors[,4] + distm_TSI_geo_dist_pcs$vectors[,5] + distm_TSI_geo_dist_pcs$vectors[,6] + distm_TSI_geo_dist_pcs$vectors[,7])

set.seed(5)

distm_genetic_geo <- anova.cca(distm_draft_TSI_genetic_geographic_with_pcs, by = "margin", permutations = 9999)
#the same result as above

p.adjust(distm_genetic_geo$`Pr(>F)`, method = "BH")

#PCs 1 and 2 are suitable to condition the model. 

#Not used - overall model testing all vectors with Fst:

fst_geo_dist <- capscale(dist_Fst_TSI_full ~ distm_TSI_geo_dist_pcs$vectors)

set.seed(5)

fst_geo_dist_anova <- anova.cca(fst_geo_dist, by = "margin", permutations = 9999)

```

Running a dbRDA for the number of # plane landings on each island over a given year (July 2020 - July 2021)

6. Import the plane data

```{r}
plane_landings_full <- read.csv(file = "plane_landings_full.csv")

```

7. Convert plane landings into 'dist'

```{r}
plane_landings_full_dist <- dist(plane_landings_full)
str(plane_landings_full_dist)
```

8. Convert plane landings into PCs

```{r}
plane_landings_full_pcs <- pcnm(plane_landings_full_dist)
```

9. Run the dbRDA with genetic, plane landings full, conditioned by geo distance

```{r}
#this is the best model.
dbRDAplane_landings <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors + Condition(distm_TSI_geo_dist_pcs$vectors[,c(1,2)]))

set.seed(5)

anova.cca(dbRDAplane_landings, by = "margin", permutations = 9999)

#This is no where near close to significance. How about if I remove the geographic distance conditioning?

#try to get the sum of squares for just geo dist
                       
#with removing the conditioning by the geo
no_geo_dbRDAplane_landings <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors)

set.seed(5)

anova.cca(no_geo_dbRDAplane_landings, by = "margin", permutations = 9999)

#removing the conditioning by the geo results in this being highly insignificant. 

```
10. Condition the model by population size

10a: Import the island population size
```{r}
#import the island population size

island_population <- read.csv("island_population.csv")
```

11: Convert this island population to 'dist' format

```{r}
island_population_dist <- dist(island_population)
str(island_population_dist)
```

12: Convert this 'dist' format to PCs

```{r}
island_population_pcs <- pcnm(island_population_dist)

```

13. Run the Fst against all the PCs for the population size

```{r}
distm_draft_TSI_genetic_population_with_pcs <- capscale(dist_Fst_TSI_full ~ island_population_pcs$vectors[,1] + island_population_pcs$vectors[,2] + island_population_pcs$vectors[,3] + island_population_pcs$vectors[,4] + island_population_pcs$vectors[,5] + island_population_pcs$vectors[,6] + island_population_pcs$vectors[,7] + island_population_pcs$vectors[,8]+ island_population_pcs$vectors[,9] + island_population_pcs$vectors[,10])

set.seed(5)

distm_genetic_populations <- anova.cca(distm_draft_TSI_genetic_population_with_pcs, by = "margin", permutations = 9999)

distm_genetic_populations

#it looks like they're 3 PCs which are significant. 

p.adjust(distm_genetic_populations$`Pr(>F)`, method = "BH")

#with the correction, there is only two PCs which are significant. That being, PC 3 and PC 9

```

14. Run the dbRDA with genetic ~ plane landings full, conditioned by both geo distance and population size

```{r}
#this is the best model.

#anova.cca for just the geography conditions

dbRDAplane_landings_geo_pop_conditions <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors + Condition(distm_TSI_geo_dist_pcs$vectors[,c(1,2)]))

plot(dbRDAplane_landings_geo_pop_conditions)
screeplot(dbRDAplane_landings_geo_pop_conditions)


set.seed(5)

anova.cca(dbRDAplane_landings_geo_pop_conditions, by = "margin", permutations = 9999)

#model with just the population conditioning

dbRDAplane_landings_pop_conditions <- capscale(dist_Fst_TSI_full ~ plane_landings_full_pcs$vectors + Condition(island_population_pcs$vectors[,c(3,9)]))

set.seed(5)

anova.cca(dbRDAplane_landings_pop_conditions, by = "margin", permutations = 9999)

```

Run the dbRDA with the boat data

1. Import the seven island boat data

```{r}
seven_islands_boat <- read.csv("boat_data_seven_islands.csv", header = T)
str(seven_islands_boat)
```

1b convert the boat data to dist format

```{r}
seven_islands_boat_dist <- dist(seven_islands_boat)
str(seven_islands_boat_dist)
  
```

1c converting the boat data to PCs

```{r}
seven_islands_boat_dist_pc <- pcnm(seven_islands_boat_dist)

#this generated PCs 1 - 5
```


2. Filter the Fst for the 7 islands only (in the correct order)

```{r}
#fst is for the 7 islands
#needs to be in the correct order.

Fst_seven_islands <- read.csv("FST_seven_islands.csv", header = T)

```

3. Convert the Fst to 'dist' format

```{r}
dist_Fst_seven_islands <- dist(Fst_seven_islands)
str(dist_Fst_TSI_full)

```


4. Filter the geo data for the 7 islands only (correct order)

```{r}

geo_seven_islands <- read.delim("geo_seven_islands.txt", header = F)

str(geo_seven_islands)

row.names(geo_seven_islands) <- geo_seven_islands$V1
geo_seven_islands_rows <- geo_seven_islands[,-1]

#column binding the coordinates together

geo_seven_islands_rows <- cbind(geo_seven_islands_rows$V2, geo_seven_islands_rows$V3)

str(geo_seven_islands_rows)

```

5. Convert filtered geo data into PCs

```{r}
#calculating the distance between points - regular distance
dist_geo_seven_islands <- distm(geo_seven_islands_rows)
str(dist_geo_seven_islands)
plot(dist_geo_seven_islands)

#now converting this to 'dist' format
dist_dist_geo_seven_islands <- dist(dist_geo_seven_islands)
  str(dist_dist_geo_seven_islands)

```

6. Converting filtered geo to PCs

```{r}
dist_geo_seven_islands_pc <- pcnm(dist_dist_geo_seven_islands)
str(dist_geo_seven_islands_pc)
#this converted to 3 PCs.
```

7. Importing the population size for each island

7a: Import the island population size
```{r}
#import the island population size as a matrix

seven_island_population <- read.csv("seven_island_population.csv")
```

11: Convert this island population to 'dist' format

```{r}
seven_island_population_dist <- dist(seven_island_population)
str(seven_island_population_dist)
```

12: Convert this 'dist' format to PCs

```{r}
seven_island_population_dist_pcs <- pcnm(seven_island_population_dist)

```

11. Filter the Fst for the 7 islands only (in the correct order)

```{r}
#fst is for the 7 islands
#needs to be in the correct order.

Fst_seven_islands <- read.csv("FST_seven_islands.csv", header = T)

```

11b. Convert the Fst to 'dist' format

```{r}
dist_Fst_seven_islands <- dist(Fst_seven_islands)
str(dist_Fst_TSI_full)

```

13. Run the Fst against all the PCs for the population size

```{r}
seven_islands_TSI_genetic_population_with_pcs <- capscale(dist_Fst_seven_islands ~ seven_island_population_dist_pcs$vectors[,1] + seven_island_population_dist_pcs$vectors[,2] + seven_island_population_dist_pcs$vectors[,3] + seven_island_population_dist_pcs$vectors[,4] + seven_island_population_dist_pcs$vectors[,5])

set.seed(5)

distm_genetic_populations <- anova.cca(seven_islands_TSI_genetic_population_with_pcs, by = "margin", permutations = 9999)

distm_genetic_populations

#it looks like there is no effect of population size.
```



. Running the dbRDA for the boat data: Preliminary model Fst - geo distance

```{r}

prelim_dbRDAboat_Fst_geo <- capscale(dist_Fst_seven_islands ~ dist_geo_seven_islands_pc$vectors[,1] + dist_geo_seven_islands_pc$vectors[,2] + dist_geo_seven_islands_pc$vectors[,3] + dist_geo_seven_islands_pc$vectors[,4] + dist_geo_seven_islands_pc$vectors[,5])

set.seed(5)

boat_genetic_geo_sig_pcs <- anova.cca(prelim_dbRDAboat_Fst_geo, by = "margin", permutations = 9999)
#the same result as above

p.adjust(boat_genetic_geo_sig_pcs$`Pr(>F)`, method = "BH")

#in this case, all of the PCs are highly insignificant. Still a result
```

8. Running the dbRDA with the boat data(i) not conditioned by geographic distance (ii) conditioned by geographic distance (iii) Conditioned only by population size.

```{r}
#genetic is: dist_Fst_seven_islands
#geo is: dist_geo_seven_islands_pc$vectors
#boat is: seven_islands_boat_dist_pc$vectors

#try without conditioning the model. Looks like Schmidt used both with and w/o geo conditioning.

dbrda_boatfull <- capscale(dist_Fst_seven_islands ~ seven_islands_boat_dist_pc$vectors)
  
genetic_boat <- anova.cca(dbrda_boatfull, by = "margin", permutations = 9999)
#the same result as above

genetic_boat

p.adjust(genetic_boat$`Pr(>F)`, method = "BH")

#conditioned by geographic distance

dbrda_boatfull_conditioned_geo <- capscale(dist_Fst_seven_islands ~ seven_islands_boat_dist_pc$vectors + Condition(dist_geo_seven_islands_pc$vectors[,c(1,2)]))

anova.cca(dbrda_boatfull_conditioned_geo, by = "margin", permutations = 9999)

#conditioned with vectors from all islands

#Conditioned only by population size - this is not significant at all.

dbrda_boatfull_condition_population <- capscale(dist_Fst_seven_islands ~ seven_island_population_dist_pcs$vectors)

anova.cca(dbrda_boatfull_condition_population, by = "margin", permutations = 9999)
```


